\doctitle

### 2023-04-11 msr3-theory.tex
Add appendices; connect MSR3 theory and practice definitions; compiles with NO WARNINGS

\subsection{Relaxation of the mixed-effects variable selection model}
\label{sec:relax}

Our strategy for obtaining approximate solutions to the
mixed-effects variable selection problem \eqref{eq:lme_loss_original_in_x} is
motivated by the sparse relaxed regularization regression (SR3) strategy
developed in \cite{Zheng2019SR3}. That is, $0\le\gam$ is relaxed
using the perspective of the negative log, i.e %log-barrier function
$\map{\vphi}{\R^q\times\R}{\R\cup\{\infty\}}$ given by
\vphi(\gam,\mu):=
\begin{cases}
-\mu\sum_{i=1}^q\ln (\gam_i/\mu)&,\ \mu>0,\\
\del_{\R^q_+}(\gam)&,\ \mu=0,\\
+\infty&,\ \mu<0.
\end{cases}
The mapping $\vphi$ is known to be a closed proper convex function and, 
for $\mu>0$, it is
essentially equivalent to the well-known log-barrier function in \eqref{eq:log_barrier_relaxation_for_ip}. 
%since in this case
%\phi_\mu(\gam)=q\mu\ln(\mu)-\mu\sum_{i=1}^q\ln (\gam_i).
For more information on the perspective mapping, its calculus,
and perspective duality, we refer the reader to \cite{ABDFM18,ABF13}.
%\red{Should we mention the perspective of $\psi(\gam)=-\ln (\gam)$, or perhaps the closure of the epi-limit? The connection to the perspective map might be useful since
%we get the joint convexity in $\mu$ and $\gam$.}
We call $\eta$ the coupling parameter and $\mu$ the log-barrier parameter
and write $\phimu(\cdot):=\vphi(\cdot,\mu)$.
%The parameter $0\le\blam$ is fixed and is related to the weak convexity of $\LL$
%discussed in Lemma \ref{lem:LL weak cvx}.
%}
The relaxed problem employs 
auxiliary variables $\tbg$ 
and relaxation parameters $0\le\eta$ and $0\le \mu$
to obtain the problem 
\begin{equation}\label{eq:relax} 
\begin{aligned}%\eq{\label{eq:lme_loss_relaxed}
%\nu_\emu:=	
\min_{\bg, \tbg} & \LL(\beta, \gamma) 
	+\phi_\mu(\gam)+\kappa_\eta(\beta-\tbeta,\gam-\tgam)
%\frac{\eta}{2}\norm{\beta - \tbeta}^2 + \frac{\blam+\eta}{2}\norm{\gamma - \tgamma}^2  
+ R(\tbeta, \tgamma) 
\\
\text{s.t. } & \tgamma \geq 0\ .
We rewrite \eqref{eq:relax} %\eqref{eq:lme_loss_relaxed} 
so as to separate the smooth and nonsmooth components to obtain
\eq{ \label{eq:relax2}
    \min_{\bg, \tbg} & \LL_{\emu}(\bg,\tbg) + R\tbg +\del_{\R^q_+}(\tgam)\ ,
%    \\
%    \text{s.t. } & 0\le\tgam,
}
%$x=(\beta, \gamma)$, 
%$w=(\tbeta, \tgamma)$, $\CC=\R^p\times\R^q_+$, and
%$\LL_\eta$ is defined in \eqref{eq:defn L sub eta}.
\eq{\label{eq:LLlam}
    \LL_{\emu}(\bg, \tbg) := 
    \LL\bg + \phimu(\gam)+\keta(\beta - \tbeta, \gamma - \tgamma).
}
Observe that, for all $\mu,\eta\in\R_+$, $\LL_{\emu},\ \nabla \LL_{\emu}$ and $\nabla^2 \LL_{\emu}$ are continuous
on $(\Rp\times\dom(\phimu))\times(\Rp\times\Rq)$ 
(see Appendix \ref{appendix:derivatives_of_lmm}) 
so that $\LL_{\emu}$ is smooth on its domain.
As in \cite{Zheng2019SR3}, we use the 
decoupling to write \eqref{eq:relax2}
as an iterated optimization problem over the smooth components of the objective.
This yields a representation of the form
\begin{equation}
    \label{eq:relax3}
    \min_{\tbg} \uemu\tbg + R\tbg+\del_{\Rqp}(\tgam),
where %$w=(\tbeta,\tgam)$ and
\begin{equation}
    \label{eq:value_function_definition}
    \uemu\tbg := \min_{\bg } 
    \LL_{\emu}(\bg, \tbg)\ .
%    \hLL(\bg,\tgam)+ %\del_{\Rqp}(\gam)+
%    \frac{\eta}{2}\norm{\bg-\tbg}^2
    %\LL(x) + \lambda\|x - w\|_2^2 
This is the formulation of the mixed-effects variable selection problem we study.
Our focus is on the optimal value function $\uemu$ which captures global 
variational information about the function $\LL$ over its domain.
We show that $\uemu$ has a locally Lipschitz continuous gradient
and that the evaluation of $\uemu$ and $\nabla \uemu$ is accomplished
by optimizing a well conditioned strongly convex function. This
allows us to apply the PGD algorithm to the function $\uemu$ rather than
the function $\LL$. Our numerical studies show that the global information 
captured by $\uemu$ significantly improves both the accuracy of the solution
obtained and the overall numerical efficiency of the algorithm.
%gives the problem structure used in the design of our
%numerical solution method. 
%Specifically, we design a 
%(PGD) %prox gradient 
%method 
%for solving this problem (see Section \ref{sec:pgd}). For this we must show that the function $u_\emu$
%is continuously differentiable with Lipschitz continuous gradient.

%\section{Prox Gradient Descent}
\subsection{Proximal Gradient Descent for $\uemu\tbg + R\tbg+\del_{\Rqp}(\tgam)$} 
\label{sec:pgd}

@@

\begin{algorithm}[H]
\SetAlgoLined
$\tw = \tw^0$, $\gammax>\tgam^0$, $\alpha < \frac{2}{L}$ 
for $L>0$.\\
 \While{not converged and $\tgam\le\gammax$}{
    $\tw^+ = \prox_{\alpha \tR }(\tw - \alpha \nabla \uemu (\tw))$\\
 }
\caption{\label{alg:MSR3}Proximal Gradient Descent for 
$\Phi_\emu$}
\end{algorithm}
\medskip

@@

%\begin{figure}[H]
%    \centering
%    \includegraphics[width=\textwidth]{Images/intuition_current.pdf}
%    \caption{Proximal Gradient Descent (PGD) takes an order of magnitude fewer iterations to converge when optimizing a value function instead of the original likelihood, because SR3-relaxation improves the conditioning of the problem, which results in a more spherical level-sets. This effect persists for both quadratic convex (first row) and non-convex components (second row) of the likelihood. The effect is stronger for  minima near the boundary of the constraint region 
%    (bottom-left panel).}
%    \label{fig:geometric_intuition_sr3}
%\end{figure}

\subsection{The Smoothness of $u_\emu$}\label{sec:smoothness}
We investigate the relationship between the problems \eqref{eq:lme_loss_original_in_x}
and \eqref{eq:relax2}, the existence of solutions to 
 \eqref{eq:relax2}, and the properties of the function $\uemu$ and its derivative.
%We begin with an examination of the function $\LL$
%and 
%The first step is to understand the role of the parameter $\blam$.

@@

\begin{lemma}[$\LL+\phimu$ is Weakly Convex]\label{lem:LL weak cvx}
Let $\LL$ be as given in \eqref{eq:lme_gam}. Then
\begin{equation}\label{eq:hess LL}
\nabla^2\LL{(\beta,\gam)}=\sum_{i=1}^m
S_i^T\begin{bmatrix}X_i^T\\ -Z_i^T\end{bmatrix}
\Omega_i(\gam)^{-1}
\begin{bmatrix}X_i& -Z_i\end{bmatrix}S_i
-\begin{bmatrix}
0&0\\ 0& \half(Z_i^T\Omega_i(\gam)^{-1}Z_i)^{\circ2}
\end{bmatrix},
\end{equation}
for all
$(\beta,\gam)\in\R^p\times\R^q_+$, where 
\[
S_i:=\begin{bmatrix}
I_q&0\\ 0&\Diag{Z_i^T\Omega_i^{-1}(X_i\beta-Y_i)}
\end{bmatrix}
\]
and, for any $A\in\R^{t\times t}$, $A^{\circ 2}:=A\circ A$. In particular, this implies
that the matrix
\begin{equation}\label{eq:psd for LL}
\begin{bmatrix}
\nabla_{\beta\beta}\LL(\beta,\gam)&\nabla_{\gam\beta}\LL(\beta,\gam)\\
\nabla_{\beta\gam}\LL(\beta,\gam)&\nabla_{\gam\gam}\LL(\beta,\gam)
+\blam I\end{bmatrix}
\end{equation}
is positive semidefinite
for $\bar\eta= \nu m$, where
\[
\nu:=max\lset{(1/2) \mu_\mmin(\Lam_i)^{-2}\sig^4_\mmax(Z_i)}{i=1,\dots,m},
\]
$\mu_\mmin(\Lam_i)$ is the smallest eigenvalue of $\Lam_i$, and
 $\sig_\mmax(Z_i)$ is the largest singular value of $Z_i,\, i=1,\dots,m$.
Consequently, for any $\tbg\in\dom{\tR}$ and $\mu\ge 0$, the mapping 
\(
\bg\mapsto \LL_{\emu}(\bg, \tbg)
%\hLL(\bg,\tgam) %:=\LL(\beta,\gam)+\frac{\blam}{2}\norm{\gam-\hgam}^2
%\sum_{i=1}^m\frac{\nu_i}{2}\norm{\gam-\hat\gam}^2,
\)
is convex for all $\eta\ge\bar\eta:=\nu m$.
In particular, this implies that $\LL+\phimu$ is weakly convex for any $\mu\ge 0$, 
and the mapping
\(
%\begin{aligned}
\bg\mapsto\LL_{\emu}(\bg,\tbg)
%&=
%\LL\bg+\phimu(\gam)+
%\keta(\beta-\tbeta,\beta-\tbeta)
%\\ &=
%\hLL(\bg,\tgam)+    \frac{\eta}{2}\norm{\bg-\tbg}^2 
%\end{aligned}
\)
is strongly convex for $\eta> \bar\eta$ with modulus of strong convexity $(\eta-\bar\eta)$
regardless of the choice of $\tbg\in\Rp\times\Rq$.
%Moreover, the mapping 
%\eq{\label{eq:defn L sub eta}
%(\beta, \gamma)\mapsto    
%\LL_{\eta}(x, w) := \tLL(x, w) + \frac{\eta}{2}\norm{x-w}^2 
%}
%is strongly convex for every $\eta>0$ and any choice of 
%$(\tbeta,\tgam)\in\R^p\times\R^q$.
\end{lemma}
\begin{proof}
The formula for $\nabla^2\LL$ is given in Appendix \ref{appendix:derivatives_of_lmm}
(see \eqref{eq:hess LL}).
By \cite[Theorem 3.1]{ABBP2021}, 
$\mu_\mmax(\left(Z_i^T\Omega_i(\gam)^{-1}Z\right)\le
\lam_\mmin^{-1}\sig_\mmax^2(Z_i)$, and since 
$\mu_\mmax(H^{\circ 2})\le \mu_\mmax^2(H)$
for all $H\in\bS^q_+$ \cite{HJ85}, we have
\[
\mu_\mmax\left(\half(Z_i^T\Omega_i(\gam)^{-1}Z_i)^{\circ2}\right)
\le (1/2) \lam_\mmin^{-2}
\sig^4_\mmax(Z_i)=:\nu_i
\qquad  i=1,\dots,m.
\]
This establishes that the matrix in \eqref{eq:psd for LL} is positive semidefinite.
Since $\phimu$ is convex, the mapping
\[
(\beta,\gam)\mapsto \LL_{\emu}(\bg,\tbg) %\hLL((\beta,\gam),\tgam)
\]
is strongly convex for any choice
of $\blam> \nu m$, where 
\eq{\label{eq:nu}
\nu:=\max_{i=1,\dots,m}\nu_i.
}
%The strong convexity statement follows from 
%the fact that
%$\nabla^2 \LL_{\eta}((\beta, \gamma), (\tbeta, \tgamma))
%=\nabla^2 \LL(\beta,\gam)+\eta I$ where $\nabla^2 \tLL(\beta,\gam)$
%is positive semidfinite.
\end{proof}
For the remainder of the paper, we assume that 
\eq{\label{eq:blam1}\eta> \nu m=:\bar\eta} 
so that the mapping
$(\beta,\gam)\mapsto \LL_{\emu}(\bg,\tbg)$ %\hLL((\beta,\gam),\tgam)$
is strongly convex with 
%$\nabla_{\bg\bg}^2\hLL(\bg,\tgam)$ 
positive definite Hessian
regardless of the 
choice of $\tgam\in\R^q$. With this in mind, the function $\uemu$ defined by
\eqref{eq:value_function_definition} resembles a Moreau envelope.
%of the convex function $\hLL+\phi_\mu$.
However, this is misleading since, in particular,
we are not even assured of the existence of solutions to
the optimization problem defining $\uemu$.

@@

Next we fix $\mu\ge 0$ and show that as 
$\eta\uparrow\infty$ the solutions to 
\eqref{eq:relax2} converge to solutions of
\begin{equation}\label{eq:log-barrier problem}
\min_{\bg\in\CC}\LL\bg+\phimu(\gam)+R\bg.
\end{equation}
In particular, for $\mu=0$, they converge to solutions of 
\eqref{eq:vs_lme}.

\begin{theorem}[Consistency as $\eta\rightarrow\infty$]\label{thm:eta consistency}
Let $\LL$ and $R$ be as in Theorem \ref{thm:relaxed existence}
and fix $\mu\ge 0$.
Let $\{\eta_k\}\subset\R_{++}$ be such that $\eta_k<\eta_{k+1}$
with $\eta_k\uparrow\infty$, and let $(\bgk,\tbgk)$ be an optimal solution to 
\eqref{eq:relax2} for $(\emu)=(\eta_k,\mu)$, $k\in\N$.
Then any limit point (equivalently, cluster point) 
$(\bbg,\hbg)$ of $\{(\bgk,\tbgk)\}$ satisfies
$\bbg=\hbg$ with $\bbg$ being an optimal solution to 
\eqref{eq:log-barrier problem}.
\end{theorem}
\begin{proof}
With no loss in generality $\eta_k>\bar\eta$ for all $k$.
Set 
\[\left.\begin{aligned}
a_k(x,w)&:=\LL_{\eta_k,\mu}(x,w)+R(w)\\ 
b_k(x,w)&:=\LL(x)+\phimu(\gam)+R(w)\\
c_k(x,w)&:=\kappa_{\eta_k}(\beta-\tbeta,\gam-\tgam)
\end{aligned}\right\}\quad \forall k\in\N,
\] 
where 
$x=(\beta,\gam)$ and $w=(\tbeta,\tgam)$ with $\kappa_\eta$ defined in
\eqref{eq:kappa}. Set $x^k=\bgk,\ \bx=\bbg,\ w^k=\tbgk$ and $\hw=\hbg$.
By Lemma \ref{lem:lb plus 1c} and Theorem \ref{thm:basic existence2} 
with $\hR=\phimu+R$, 
%\ref{thm:relaxed existence}, 
there is an optimal solution $\xmu$ 
to \eqref{eq:log-barrier problem} %\eqref{eq:vs_lme} 
yielding an optimal value of $\vmu$ for which 
$a_k(x^k,w^k)\le a_k(\xmu,\xmu)=\vmu$ for all $k\in\N$. Hence, the sequence
$\{a_k(x^k,w^k)\}$ is upper bounded by $\vmu$.
Since
\[
a_k(x^k,w^k)\le a_k(x^{k+1},w^{k+1})\le a_{k+1}(x^{k+1},w^{k+1}),
\]
there exists $\tv$ such that
$a_k(x^k,w^k)\uparrow\tv\le \vmu$.
Next, observe that
\[
a_k(x^k,w^k) \le a_k(x^{k+1},w^{k+1})\ \ \text{ and }\ \ 
a_{k+1}(x^{k+1},w^{k+1})\le a_{k+1}(x^{k},w^{k}).
\]
By adding these inequalities together we find that 
$\norm{x^{k+1}-w^{k+1}}\le \norm{x^{k}-w^{k}}$ so that 
$\norm{x^{k}-w^{k}}\downarrow\tkappa$ for some $\tkappa\ge 0$.
We also have
\[\begin{aligned}
b_k(x^k,w^k)+(\eta_k/2)\norm{x^{k}-w^{k}}
&=a_k(x^k,w^k)
\\ &\le a_k(x^{k+1},w^{k+1})
\\ &=b_{k+1}(x^{k+1},w^{k+1})+(\eta_k/2)\norm{x^{k+1}-w^{k+1}}
\\ &\le b_{k+1}(x^{k+1},w^{k+1})+(\eta_k/2)\norm{x^{k}-w^{k}},
\end{aligned}\]
which gives $b_k(x^k,w^k)\le b_{k+1}(x^{k+1},w^{k+1})\le \tv$.
Therefore, $b_k(x^k,w^k)\uparrow \hv$ for some $\hv\le\tv$.
Consequently,
\[
\tkappa=\lim_k\norm{x^{k}-w^{k}}=
\lim_k\eta_k^{-1}[a_k(x^k,w^k)-b_k(x^k,w^k)]=0.
\]
Therefore, if $(\bx,\bw)$ is any limit point of the sequence $\{(x^k,w^k)\}$,
then $\bx=\bw$ and $\LL(\bx)+\phimu(\bgam)+R(\bx)=\vmu$ 
since $\LL(x^k)+\phimu(\gam^k)+R(w^k)\le a_k(x^k,w^k)\le \vmu$
for all $k\in\N$.
\end{proof}
We now pair Theorem \ref{thm:eta consistency} with a consistency result for the barrier parameter $\mu$.

>>>>

%gives the problem structure used in the design of our
%numerical solution method. 
%Specifically, we design a 
%(PGD) %prox gradient 
%method 
%for solving this problem (see Section \ref{sec:pgd}). For this we must show that the function $u_\emu$
%is continuously differentiable with Lipschitz continuous gradient.

%\section{Prox Gradient Descent}
\subsection{Proximal Gradient Descent for \texorpdfstring{$\uemu\tbg + R\tbg+\del_{\Rqp}(\tgam)$}{}} 
\label{sec:pgd}

@@

@@

%\begin{figure}[t]
%    \centering
%    \includegraphics[width=\textwidth]{Images/intuition_current.pdf}
%    \caption{Proximal Gradient Descent (PGD) takes an order of magnitude fewer iterations to converge when optimizing a value function instead of the original likelihood, because SR3-relaxation improves the conditioning of the problem, which results in a more spherical level-sets. This effect persists for both quadratic convex (first row) and non-convex components (second row) of the likelihood. The effect is stronger for  minima near the boundary of the constraint region 
%    (bottom-left panel).}
%    \label{fig:geometric_intuition_sr3}
%\end{figure}

\subsection{The Smoothness of \texorpdfstring{$u_\emu$}{}}\label{sec:smoothness}
We investigate the relationship between the problems \eqref{eq:lme_loss_original_in_x}
and \eqref{eq:relax2}, the existence of solutions to 
 \eqref{eq:relax2}, and the properties of the function $\uemu$ and its derivative.
%We begin with an examination of the function $\LL$
%and 
%The first step is to understand the role of the parameter $\blam$.

@@

\begin{lemma}[$\LL+\phimu$ is Weakly Convex]\label{lem:LL weak cvx}
Let $\LL$ be as given in \eqref{eq:lme_diagonal_setup}. Then
\begin{equation}\label{eq:hess LL}
\nabla^2\LL{(\beta,\gam)}=\sum_{i=1}^m
S_i^T\begin{bmatrix}X_i^T\\ -Z_i^T\end{bmatrix}
\Omega_i(\gam)^{-1}
\begin{bmatrix}X_i& -Z_i\end{bmatrix}S_i
-\begin{bmatrix}
0&0\\ 0& \half(Z_i^T\Omega_i(\gam)^{-1}Z_i)^{\circ2}
\end{bmatrix},
\end{equation}
for all
$(\beta,\gam)\in\R^p\times\R^q_+$, where 
\[
S_i:=\begin{bmatrix}
I_q&0\\ 0&\Diag{Z_i^T\Omega_i^{-1}(X_i\beta-Y_i)}
\end{bmatrix}
\]
and, for any $A\in\R^{t\times t}$, $A^{\circ 2}:=A\circ A$. In particular, this implies
that the matrix
\begin{equation}\label{eq:psd for LL}
\begin{bmatrix}
\nabla_{\beta\beta}\LL(\beta,\gam)&\nabla_{\gam\beta}\LL(\beta,\gam)\\
\nabla_{\beta\gam}\LL(\beta,\gam)&\nabla_{\gam\gam}\LL(\beta,\gam)
+\blam I\end{bmatrix}
\end{equation}
is positive semidefinite
for $\bar\eta= \nu m$, where
\[
\nu:=max\lset{(1/2) \mu_\mmin(\Lam_i)^{-2}\sig^4_\mmax(Z_i)}{i=1,\dots,m},
\]
$\mu_\mmin(\Lam_i)$ is the smallest eigenvalue of $\Lam_i$, and
 $\sig_\mmax(Z_i)$ is the largest singular value of $Z_i,\, i=1,\dots,m$.
Consequently, for any $\tbg\in\dom{\tR}$ and $\mu\ge 0$, the mapping 
\(
\bg\mapsto \LL_{\emu}(\bg, \tbg)
%\hLL(\bg,\tgam) %:=\LL(\beta,\gam)+\frac{\blam}{2}\norm{\gam-\hgam}^2
%\sum_{i=1}^m\frac{\nu_i}{2}\norm{\gam-\hat\gam}^2,
\)
is convex for all $\eta\ge\bar\eta:=\nu m$.
In particular, this implies that $\LL+\phimu$ is weakly convex for any $\mu\ge 0$, 
and the mapping
\(
%\begin{aligned}
\bg\mapsto\LL_{\emu}(\bg,\tbg)
%&=
%\LL\bg+\phimu(\gam)+
%\keta(\beta-\tbeta,\beta-\tbeta)
%\\ &=
%\hLL(\bg,\tgam)+    \frac{\eta}{2}\norm{\bg-\tbg}^2 
%\end{aligned}
\)
is strongly convex for $\eta> \bar\eta$ with modulus of strong convexity $(\eta-\bar\eta)$
regardless of the choice of $\tbg\in\Rp\times\Rq$.
%Moreover, the mapping 
%\eq{\label{eq:defn L sub eta}
%(\beta, \gamma)\mapsto    
%\LL_{\eta}(x, w) := \tLL(x, w) + \frac{\eta}{2}\norm{x-w}^2 
%}
%is strongly convex for every $\eta>0$ and any choice of 
%$(\tbeta,\tgam)\in\R^p\times\R^q$.
\end{lemma}
\begin{proof}
The formula for $\nabla^2\LL$ is given in Appendix \ref{appendix:derivatives_of_lmm}
(see \eqref{eq:hess LL}).
By \cite[Theorem 3.1]{ABBP2021}, 
$\mu_\mmax(\left(Z_i^T\Omega_i(\gam)^{-1}Z\right)\le
\lam_\mmin^{-1}\sig_\mmax^2(Z_i)$, and since 
$\mu_\mmax(H^{\circ 2})\le \mu_\mmax^2(H)$
for all $H\in\bS^q_+$ \cite{HJ85}, we have
\[
\mu_\mmax\left(\half(Z_i^T\Omega_i(\gam)^{-1}Z_i)^{\circ2}\right)
\le (1/2) \lam_\mmin^{-2}
\sig^4_\mmax(Z_i)=:\nu_i
\qquad  i=1,\dots,m.
\]
This establishes that the matrix in \eqref{eq:psd for LL} is positive semidefinite.
Since $\phimu$ is convex, the mapping
\[
(\beta,\gam)\mapsto \LL_{\emu}(\bg,\tbg) %\hLL((\beta,\gam),\tgam)
\]
is strongly convex for any choice
of $\blam> \nu m$, where 
\eq{\label{eq:nu}
\nu:=\max_{i=1,\dots,m}\nu_i.
}
%The strong convexity statement follows from 
%the fact that
%$\nabla^2 \LL_{\eta}((\beta, \gamma), (\tbeta, \tgamma))
%=\nabla^2 \LL(\beta,\gam)+\eta I$ where $\nabla^2 \tLL(\beta,\gam)$
%is positive semidfinite.
\end{proof}
For the remainder of the paper, we assume that 
\eq{\label{eq:blam1}\eta> \nu m=:\bar\eta} 
so that the mapping
$(\beta,\gam)\mapsto \LL_{\emu}(\bg,\tbg)$ %\hLL((\beta,\gam),\tgam)$
is strongly convex with 
%$\nabla_{\bg\bg}^2\hLL(\bg,\tgam)$ 
positive definite Hessian
regardless of the 
choice of $\tgam\in\R^q$. With this in mind, the function $\uemu$ defined by
\eqref{eq:value_function_definition} resembles a Moreau envelope.
%of the convex function $\hLL+\phi_\mu$.
However, this is misleading since, in particular,
we are not even assured of the existence of solutions to
the optimization problem defining $\uemu$.

@@

Next we fix $\mu\ge 0$ and show that as 
$\eta\uparrow\infty$ the solutions to 
\eqref{eq:relax2} converge to solutions of
\begin{equation}\label{eq:log-barrier problem}
\min_{\bg\in\CC}\LL\bg+\phimu(\gam)+R\bg.
\end{equation}
In particular, for $\mu=0$, they converge to solutions of 
\eqref{eq:lme_loss_original_in_x}.

\begin{theorem}[Consistency as $\eta\rightarrow\infty$]\label{thm:eta consistency}
Let $\LL$ and $R$ be as in Theorem \ref{thm:relaxed existence}
and fix $\mu\ge 0$.
Let $\{\eta_k\}\subset\R_{++}$ be such that $\eta_k<\eta_{k+1}$
with $\eta_k\uparrow\infty$, and let $(\bgk,\tbgk)$ be an optimal solution to 
\eqref{eq:relax2} for $(\emu)=(\eta_k,\mu)$, $k\in\N$.
Then any limit point (equivalently, cluster point) 
$(\bbg,\hbg)$ of $\{(\bgk,\tbgk)\}$ satisfies
$\bbg=\hbg$ with $\bbg$ being an optimal solution to 
\eqref{eq:log-barrier problem}.
\end{theorem}
\begin{proof}
With no loss in generality $\eta_k>\bar\eta$ for all $k$.
Set 
\[\left.\begin{aligned}
a_k(x,w)&:=\LL_{\eta_k,\mu}(x,w)+R(w)\\ 
b_k(x,w)&:=\LL(x)+\phimu(\gam)+R(w)\\
c_k(x,w)&:=\kappa_{\eta_k}(\beta-\tbeta,\gam-\tgam)
\end{aligned}\right\}\quad \forall k\in\N,
\] 
where 
$x=(\beta,\gam)$ and $w=(\tbeta,\tgam)$ with $\kappa_\eta$ defined in
\eqref{eq:kappa}. Set $x^k=\bgk,\ \bx=\bbg,\ w^k=\tbgk$ and $\hw=\hbg$.
By Lemma \ref{lem:lb plus 1c} and Theorem \ref{thm:basic existence2} 
with $\hR=\phimu+R$, 
%\ref{thm:relaxed existence}, 
there is an optimal solution $\xmu$ 
to \eqref{eq:log-barrier problem} %\eqref{eq:lme_loss_original_in_x} 
yielding an optimal value of $\vmu$ for which 
$a_k(x^k,w^k)\le a_k(\xmu,\xmu)=\vmu$ for all $k\in\N$. Hence, the sequence
$\{a_k(x^k,w^k)\}$ is upper bounded by $\vmu$.
Since
\[
a_k(x^k,w^k)\le a_k(x^{k+1},w^{k+1})\le a_{k+1}(x^{k+1},w^{k+1}),
\]
there exists $\tv$ such that
$a_k(x^k,w^k)\uparrow\tv\le \vmu$.
Next, observe that
\[
a_k(x^k,w^k) \le a_k(x^{k+1},w^{k+1})\ \ \text{ and }\ \ 
a_{k+1}(x^{k+1},w^{k+1})\le a_{k+1}(x^{k},w^{k}).
\]
By adding these inequalities together we find that 
$\norm{x^{k+1}-w^{k+1}}\le \norm{x^{k}-w^{k}}$ so that 
$\norm{x^{k}-w^{k}}\downarrow\tkappa$ for some $\tkappa\ge 0$.
We also have
\[\begin{aligned}
b_k(x^k,w^k)+(\eta_k/2)\norm{x^{k}-w^{k}}
&=a_k(x^k,w^k)
\\ &\le a_k(x^{k+1},w^{k+1})
\\ &=b_{k+1}(x^{k+1},w^{k+1})+(\eta_k/2)\norm{x^{k+1}-w^{k+1}}
\\ &\le b_{k+1}(x^{k+1},w^{k+1})+(\eta_k/2)\norm{x^{k}-w^{k}},
\end{aligned}\]
which gives $b_k(x^k,w^k)\le b_{k+1}(x^{k+1},w^{k+1})\le \tv$.
Therefore, $b_k(x^k,w^k)\uparrow \hv$ for some $\hv\le\tv$.
Consequently,
\[
\tkappa=\lim_k\norm{x^{k}-w^{k}}=
\lim_k\eta_k^{-1}[a_k(x^k,w^k)-b_k(x^k,w^k)]=0.
\]
Therefore, if $(\bx,\bw)$ is any limit point of the sequence $\{(x^k,w^k)\}$,
then $\bx=\bw$ and $\LL(\bx)+\phimu(\bgam)+R(\bx)=\vmu$ 
since $\LL(x^k)+\phimu(\gam^k)+R(w^k)\le a_k(x^k,w^k)\le \vmu$
for all $k\in\N$.
\end{proof}
We now pair Theorem \ref{thm:eta consistency} with a consistency result for the barrier parameter $\mu$.

\begin{theorem}[Consistency as $\mu\rightarrow 0$]
\label{thm:mu consistency}
Let $\LL$ and $R$ be as in Theorem \ref{thm:relaxed existence}. 
For every $\mu\ge 0$, problem 
\eqref{eq:log-barrier problem} has a solution $(\beta_\mu,\gam_\mu)$.
Moreover, if $\{\mu_k\}\subset\R_{++}$ is such that $\mu_k\downarrow 0$,
then the sequence $\{(\beta_{\mu_k},\gam_{\mu_k})\}$ is bounded
and every limit point of the sequence %$\{(\beta_\mu,\gam_\mu):\ \mu>0\}$
%$(\bbeta,\bgam)$ 
%as $\mu\rightarrow 0$ 
is a 
solution to \eqref{eq:lme_loss_original_in_x}.
\end{theorem}
\begin{proof}
The existence of $(\beta_\mu,\gam_\mu)$ for all $\mu\ge 0$ 
follows immediately from 
Lemma \ref{lem:lb plus 1c} and Theorem \ref{thm:basic existence2} with
$\hR=R+\phimu$. 
%Let $(\bbeta,\bgam)$ be a 
%limit point of $\{(\beta_\mu,\gam_\mu):\ \mu>0\}$
%as $\mu\rightarrow 0$. 
%That is, there is a sequence 
Let $\mu_k\downarrow 0$
and set $(\beta^k,\gam^k):=(\beta_{\mu_k},\gam_{\mu_k})$.
%\rightarrow (\bbeta,\bgam)$.
Set $\tLL:=\LL+R+\del_{\Rp\times\Rqp}$ so that 
the objective in \eqref{eq:log-barrier problem} is $\tLL+\phimu$
and the objective in \eqref{eq:lme_loss_original_in_x} is $\tLL$
with $(\beta_0,\gam_0)$ a solution to \eqref{eq:lme_loss_original_in_x}
by definition. 
Observe that
\[\begin{aligned}
\tLL\bgk+\phi_{\mu_k}(\gam^k)&\le 
\tLL(\beta^{k+1},\gam^{k+1})+\phi_{\mu_k}(\gam^{k+1})\quad \text{and}
\\
\tLL(\beta^{k+1},\gam^{k+1})+\phi_{\mu_{k+1}}(\gam^{k+1})&\le 
\tLL(\beta^{k},\gam^{k})+\phi_{\mu_{k+1}}(\gam^k)
\end{aligned}\]
Summing these inequalities yields the inquality
\[
(\mu_k-\mu_{k+1})\sum_{i=1}^q\ln(\gam^{k}_i)
\ge 
(\mu_k-\mu_{k+1})\sum_{i=1}^q\ln(\gam^{k+1}_i),
\]
so $\{\sum_{i=1}^q\ln(\gam^{k}_i)\}$ is a non-increasing sequence.
Therefore,
\[\begin{aligned}
\tLL(\beta^{k+1},\gam^{k+1})+\phi_{\mu_{k+1}}(\gam^{k+1})&\le 
\tLL(\beta^{k},\gam^{k})+\phi_{\mu_{k+1}}(\gam^k)
\\ &\le
\tLL(\beta^{k},\gam^{k})+\phi_{\mu_{k+1}}(\gam^{k+1})
\end{aligned}\]
which implies that $\{\tLL(\beta^{k},\gam^{k})\}$
is also a non-increasing sequence and bounded below by 
$\tLL(\beta_0,\gam_0)$. 
Since
Theorem \ref{thm:basic existence2} tells us that $\tLL$ is level compact,
the sequence $\{\bgk\}$ is bounded. 
Let $(\bbeta,\bgam)\in\Rp\times\Rqp$ be any
limit point of $\{\bgk\}$ and let $J\subset\N$ be such that
$\bgk\overset{J}{\rightarrow}(\bbeta,\bgam)$. Then
\[
\tLL\bgk+\phi_{\mu_k}(\gam^k)\le
\tLL\bg+\phi_{\mu_k}(\gam) \quad \forall\,\bg\in\Rp\times\R^q_{++}.
\]
Since $\tLL$ is continuous on $\Rp\times\Rqp$ and the perspective function
$\phi_{\mu}(\gam)=\varphi(\mu,\gam)$ is lsc on $\Rp\times\Rqp$, we have
\[
\tLL(\bbeta,\bgam)\le\liminf_{k\in J}(\tLL\bgk+\phi_{\mu_k}(\gam^k))\le
\tLL\bg \quad \forall\,\bg\in\Rp\times\R^q_{++}.
\]
Consequently, the continuity of $\tLL$ on $\Rp\times\Rqp$ implies
that $(\bbeta,\bgam)$ solves \eqref{eq:lme_loss_original_in_x}.
\end{proof}
